Moisture absorption and drying alter nonadditive litter flammability in a mixed conifer forest

### 2020-05-01 ms/Magalhaes_Schwilk_litter_flam_moisture.tex
fixup
Update number of words

\documentclass[letterpaper,12pt]{article}
\usepackage{caption}
\DeclareCaptionLabelFormat{adja-page}{\hrulefill\\#1 #2 \emph{(previous page)}}
\usepackage{jecol}
\usepackage[notablist,nofiglist]{endfloat}
%\usepackage{color}
%\usepackage{soul}
%\renewcommand{\hl}[1]{#1}

@@

\nwords{8200}
\ntables{4}
\nfig{5}
\nref{55}

@@

Field sites for this study were located in Sequoia and Kings Canyon National
Parks, California, USA (N$36^\circ 36'$, W$118^\circ 42'$) between 1600 and
2400\,m elevation in mixed-conifer forest. We chose eight dominant tree species
that are representative of this type of forest: \emph{Pinus jeffreyi} Grev. \&
Balf., \emph{Pinus lambertiana} Dougl., \emph{Pinus ponderosa} Dougl. ex Laws.,
\emph{Abies concolor} (Gord. \& Glend.) Lindl. ex Hildebr., \emph{Abies
  magnifica} A. Murr., \emph{Calocedrus decurrens} (Torr.) Florin,
\emph{Quercus kelloggii} Newb., and \emph{Sequoiadendron giganteum} (Lindl.) J.
Buchholz. In this study, we consider ``litter'' the top layer of leaves and small
twigs less than 1\,cm diameter that have fallen in the previous year (mostly
undecomposed or lightly decomposed litter representing fuels classified as
1-hour and 10-hour fuels). Leaf litter was collected in the summer (mid-June to
mid-July) in 2012 from 21 separate sites across the study area. We collected
from a minimum of four populations per species with each population within a
species was 9--32\,km distant from its neighbors. An exception to this was
\emph{A. magnifica} which is restricted to the higher elevations and could only
be collected from one area due to designated wilderness boundaries and
logistical constraints. Within that area, \emph{A. magnifica} litter was
collected from over 10 different individuals. For the remaining species, the
collection involved 2--4 individual trees at each site which were separated
from one another by at least 10\,m. The litter was collected 2\,m away from the
trunk to obtain a more uniform sample of leaf and fine branchlet litter,
because bark and heavier branches tend to fall closer to the tree trunk. The
top 2.5-5\,cm of litter were collected by raking and shoveling into large bags.
All litter was oven dried at $60^\circ$C to $< 5$\% fuel moisture by dry
weight, then weighed and divided into moisture trials. The collected litter was
stored in large paper bags and cardboard boxes in a lab storage space. We
minimized handling and moving the bagged litter to avoid mechanical damage and
changes to the particle size distributions. The moisture manipulations and
subsequent burning trials were conducted in April--August 2015.

@@

We performed flammability tests across multiple moisture levels on monospecific
litter beds and on mixed litter beds with each species or mixture replicated 5
times at each moisture level. We used the initial dry-down experiment to guide
our timing so as to attempt to burn samples over a range of moisture contents.
We staggered saturation and start of drying so we could conduct our burning in
a blocked design with multiple species at the same moisture content per burning
day. We aimed for moisture contents from 10\% to 100\% on a dry mass basis because
preliminary trials on damp fuels suggested that across these fuels ignition
will not occur above 100\% fuel moisture. To expand our data on the dry end of
this gradient, we supplemented these data with data from an earlier experiment
on oven-dried litter \citep{Magalhaes+Schwilk-2012}. This supplemental data
added five replicates per species at an oven dried moisture content of $<$ 5\%.
We assigned these replicates a time since wetting value of 144 hours equivalent
to the longest time since drying for the wetted samples.

@@

\begin{table}
  \caption{Linear mixed model results for moisture content as a function of
    time since wetting and species. Approximate degrees of freedom (df), pseudo F
    statistics and p-values were calculated by the Kenward-Roger approximation.
    \citep{Kenward_Roger-1997}. Model coefficients are reported in SI Table~1.}
  \label{tab:drydown}
\centering
%% \tablesize{} %% You can specify the fontsize here, e.g., \tablesize{\footnotesize}. If commented out \small will be used.
\begin{tabular}{llll}
  \toprule
Effect & df & $F$ & $P$ \\ 
  \midrule
  hour & 1, 183.73 & 1599.59 & $<$.0001 \\ 
  species & 7, 89.73 & 36.15 & $<$.0001 \\ 
  hour:species & 7, 184.15 & 23.06 & $<$.0001 \\ 
   \bottomrule
\end{tabular}
\end{table}

@@

\subsection*{Moisture dynamics and flammability across taxonomic groups}

@@

Conceptualization, DWS and RMQM; methodology, RMQM and DWS; experiments, RMQM;
coding and data analysis, DWS and RMQM; writing DWS and RMQM; writing--review
and editing, DWS

>>>>

\documentclass[letterpaper,12pt]{article}
\usepackage{jecol}
\usepackage[notablist,nofiglist]{endfloat}
\usepackage{caption}
\DeclareCaptionLabelFormat{adja-page}{\hrulefill\\#1 #2 \emph{(previous page)}}

@@

\nwords{6330 (main text only), 8000 (abstract, text, captions) }
\ntables{4}
\nfig{5}
\nref{55}

@@

Field sites for this study were located in Sequoia and Kings Canyon National
Parks, California, USA (N$36^\circ 36'$, W$118^\circ 42'$) between 1600 and
2400\,m elevation in mixed-conifer forest. We chose eight dominant tree species
that are representative of this type of forest: \emph{Pinus jeffreyi} Grev. \&
Balf., \emph{Pinus lambertiana} Dougl., \emph{Pinus ponderosa} Dougl. ex Laws.,
\emph{Abies concolor} (Gord. \& Glend.) Lindl. ex Hildebr., \emph{Abies
  magnifica} A. Murr., \emph{Calocedrus decurrens} (Torr.) Florin,
\emph{Quercus kelloggii} Newb., and \emph{Sequoiadendron giganteum} (Lindl.) J.
Buchholz. In this study, we consider ``litter'' the top layer of leaves and small
twigs less than 1\,cm diameter that have fallen in the previous year (mostly
undecomposed or lightly decomposed litter representing fuels classified as
1-hour and 10-hour fuels). Leaf litter was collected in the summer (mid-June to
mid-July) in 2012 from 21 separate sites across the study area. We collected
from a minimum of four populations per species with each population within a
species was 9--32\,km distant from its neighbors. An exception to this was
\emph{A. magnifica} which is restricted to the higher elevations and could only
be collected from one area due to designated wilderness boundaries and
logistical constraints. Within that area, \emph{A. magnifica} litter was
collected from over 10 different individuals. For the remaining species, the
collection involved 2 -- 4 individual trees at each site which were separated
from one another by at least 10\,m. The litter was collected 2\,m away from the
trunk to obtain a more uniform sample of leaf and fine branchlet litter,
because bark and heavier branches tend to fall closer to the tree trunk. The
top 2.5 -- 5\,cm of litter were collected by raking and shoveling into large bags.
All litter was oven dried at $60^\circ$C to $< 5$\% fuel moisture by dry
weight, then weighed and divided into moisture trials. The collected litter was
stored in large paper bags and cardboard boxes in a lab storage space. We
minimized handling and moving the bagged litter to avoid mechanical damage and
changes to the particle size distributions. The moisture manipulations and
subsequent burning trials were conducted in April -- August 2015.

@@

We performed flammability tests across multiple moisture levels on monospecific
litter beds and on mixed litter beds with each species or mixture replicated
five times at each moisture level. We used the initial dry-down experiment to
guide our timing so as to attempt to burn samples over a range of moisture
contents. We staggered saturation and start of drying so we could conduct our
burning in a blocked design with multiple species at the same moisture content
per burning day. We aimed for moisture contents from 10\% to 100\% on a dry
mass basis because preliminary trials on damp fuels suggested that across these
fuels ignition will not occur above 100\% fuel moisture. To expand our data on
the dry end of this gradient, we supplemented these data with data from an
earlier experiment on oven-dried litter \citep{Magalhaes+Schwilk-2012}. This
supplemental data added five replicates per species at an oven dried moisture
content of $<$ 5\%. We assigned these replicates a time since wetting value of
144 hours equivalent to the longest time since drying for the wetted samples.

@@

\begin{table}
  \caption{Linear mixed model results for moisture content as a function of
    time since wetting and species. Approximate degrees of freedom (df), pseudo F
    statistics and p-values were calculated by the Kenward-Roger approximation.
    \citep{Kenward_Roger-1997}. Model coefficients are reported in SI Table~1.
  \label{tab:drydown}}
\centering
%% \tablesize{} %% You can specify the fontsize here, e.g., \tablesize{\footnotesize}. If commented out \small will be used.
\begin{tabular}{llll}
  \toprule
Effect & df & $F$ & $P$ \\ 
  \midrule
  hour & 1, 183.73 & 1599.59 & $<$.0001 \\ 
  species & 7, 89.73 & 36.15 & $<$.0001 \\ 
  hour:species & 7, 184.15 & 23.06 & $<$.0001 \\ 
   \bottomrule
\end{tabular}
\end{table}

@@

\subsection*{Moisture dynamics and flammability}

@@

Conceptualization, DWS and RMQM; methodology, RMQM and DWS; experiments, RMQM;
coding and data analysis, DWS and RMQM; writing DWS and RMQM; writing--review
and editing, DWS.
